% Created 2021-12-23 Thu 23:06
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{looph0le}
\date{\today}
\title{System Engineering And System Analysis And Design}
\hypersetup{
 pdfauthor={looph0le},
 pdftitle={System Engineering And System Analysis And Design},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.6)}, 
 pdflang={English}}
\usepackage{biblatex}

\begin{document}

\maketitle
\tableofcontents


\section{Introduction}
\label{sec:orgb98d17f}
If we want to transfer data or information between the computers there are two ways.
\begin{enumerate}
\item Without using network
\item With the help of networks
\end{enumerate}
Just imagine that there are two computers which do not have any connection, still we want to transfer
data then we have to use extra resources like Pen Drive, CD or DVD and compy it to other computer.
This process is time consuming and requires more resources and money.
With the help of network. We can transfer data or information all over the world from one system to another system.
This process is easy, fast, secure and reliable.
\subsection{Definition of Network:}
\label{sec:org37b1f1b}
\subsubsection{Computer Networks}
\label{sec:org7890974}
Computer network is defined as two or more computers are connected in such a way that they can
share their data, information as well as their resources.
Computer Network is the interconnected collection of autonoumous computer and peripheral devices.
The links connecting the devices are called \emph{communication channels}.
The concept of connecting computers for sharing resources or data are called \emph{Networking}.
Two computer are said to be interconnected if they are able to exchange \emph{data} and \emph{information}.
\subsubsection{Characteristics of Network}
\label{sec:org828f9b7}
\begin{enumerate}
\item Resource Sharing:
\label{sec:org8fc229c}
It means all the computers within the network are able to share their resources which include
all the programs, data, and devices available to anyone on the network without considering the
physical location of the resource and user.
\item High reliability:
\label{sec:orge58ea6c}
Network provides high reliability by having alternative sources of supply.
\item Saving money:
\label{sec:org29350d5}
Mostly in networking, one computer work as server and other as client means only server computer
require connection with resource and all other client can easily access it. So, it requires less
hardware, hence saving a ton of money.
\item Powerful communication medium:
\label{sec:orgb0a1e63}
One a network, when one person make any change to a document. The other person can see the change
immediatly instead of waiting several days.
\item Scalability:
\label{sec:org7280526}
Network is able to increase system performace as the work grows, just by upgrading the hardware.
\item Low cost:
\label{sec:org58245c8}
Using networks, we are able to talk with any person at any part of the world with a low cost
compare to telephone system.
\item Time saving:
\label{sec:org3d69505}
E-mail services require less time compare to postal services.
\item Flexibility of equipment location:
\label{sec:org9257518}
We can change location anytime.
\item Improved performace:
\label{sec:orgedfec3a}
We can improve performace by adding new efficient network hardware.
\end{enumerate}
\subsubsection{Networks are used by people because:}
\label{sec:org1d02990}
\begin{enumerate}
\item Access to remote information
\label{sec:orgf2f06b6}
\item Person to person communication
\label{sec:org1fe11d5}
\item Interactive Entertaiment
\label{sec:org5433d2c}
\end{enumerate}
\subsubsection{Advantages/Needs of Computer Network:}
\label{sec:org07e17fa}
\begin{enumerate}
\item Efficient management of resources
\label{sec:org71c3d64}
\item Faster Data sharing
\label{sec:orgc71df5a}
\item Keeping information reliable and up to date
\label{sec:org68711f9}
\item High reliability
\label{sec:orgb7037c1}
\item Efficient communication
\label{sec:org1c1e376}
\end{enumerate}
\subsection{Network Criteria:}
\label{sec:org451ec03}
Effective and efficient network must be able to meet following criteria.
\subsubsection{Performace}
\label{sec:org939e43c}
Performace of a network is measured by two times.
\begin{enumerate}
\item Transit time
\label{sec:orgad8b35b}
Time requied for a message to travel from one device to anoter.
\item Response time
\label{sec:org875e993}
Elapsed time between inquiry and response.
\end{enumerate}
\subsubsection{Reliablity}
\label{sec:org56a9c6a}
Network reliability is measured by
\begin{enumerate}
\item Frequency of failure
\label{sec:orgdd186d3}
All networks fail occasionally.
A network that fails often however is a little value.
\item Recovery time of a network after a failure
\label{sec:orgda618d6}
A network that recovers quickly is more useful that one that does not.
\item Catastrophe events
\label{sec:orga5ccc70}
Network must be protected from the catastrophe events like fire, earthquake, tsunami, etc.
\item Security
\label{sec:orge484fc8}
Network security include protecting data from unauthorized access and viruses
\begin{enumerate}
\item Unautorized access
\label{sec:org6c3663e}
Sensitive data must be protected from unauthorized access.
\item Viruses
\label{sec:orgdcc3cd6}
Network is accessible from many access points and can be susceptible to computer viruses.
A virus is introduced to code that damages the system. A good network is protected to computer viruses.
\end{enumerate}
\end{enumerate}
\subsection{Network Applications:}
\label{sec:org6d80ad6}
\subsubsection{Financial Services}
\label{sec:org405b9bf}
Credit history searches, foreign exchange, investment services and electronic money transfer, etc.
\subsubsection{Marketing and sales}
\label{sec:org03d5b2d}
Marketing people use them for collecting, exchange and analyze data relating to customer.
Sales applications include teleshopping, online reservation services, railway, airline, etc.
\subsubsection{Electronic messaging}
\label{sec:org8cdd962}
E-mail services are possible using networks.
\subsubsection{Manufacturing}
\label{sec:orgf9cf8e0}
Computer networks are used in manufacturing processes.
\subsubsection{Directory service}
\label{sec:org56e8dde}
It allows list of files to be stored in particular location and search operation for the file.
\subsubsection{Information Services}
\label{sec:orgbf977f3}
World wide web site offers the information services.
\subsubsection{Cable television}
\label{sec:org3c80206}
This also uses computer networking concept
\subsubsection{Teleconferencing}
\label{sec:org00519b9}
It allows conference to occure witout the particular being in same place.
It includes text, audio and video conferencing.
\subsubsection{Electronic data exchange}
\label{sec:org3a27551}
It allows business information to be transferred witout using paper.
\subsection{Network Hardware}
\label{sec:orgeb306ae}
\end{document}
